### 2015-10-30 paper/paper.tex
started with the visualization chapter

\documentclass[journal, letterpaper]{IEEEtran}
%\documentclass{scrartcl}

\usepackage[english]{babel}
%\usepackage[latin1]{inputenc}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[pdftex]{hyperref}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmiccomment}[1]{\ \ // #1} % C-like // Comments

\hyphenation{render}

% No clubs and widows allowed
\clubpenalty10000
\widowpenalty10000
\displaywidowpenalty=10000
\DeclareMathOperator*{\argmin}{arg\,min}
\begin{document}

%\title{Simulating elastic spheres without external forces}
%\subtitle{Project 1 for class CS6491 Computer Graphics}
\title{Curve Average \\
	{\large Project 3 for class CS6491 Computer Graphics}}
%\author{Sebastian Weiss}
\author{Sebastian Weiss, Can Erdogan \\ \today}
%\date{\today}

\maketitle

\begin{tikzpicture}[remember picture,overlay]
   \node[anchor=north east,inner sep=0pt] at (current page.north east)
              {\includegraphics[scale=1.5]{pic}};
\end{tikzpicture}

\section{Objective}
Formally, for two curves $A = \{a_0 ... a_{n-1}\}$ and $B = \{b_0 ... b_{n-1}\}$, each with $n$ points and $\forall x \in (A \cup B), x \in \mathcal{R}^3$,
the goal is to find a curve $C$ such that for each point $c \in C$, and its closest projections
\begin{align}
a^c &= \operatornamewithlimits{argmin}\limits_{a \in A} ||a-c|| \text{, and} \\
b^c &= \operatornamewithlimits{argmin}\limits_{b \in B} ||b-c||
\end{align}

\noindent the following holds true:

\begin{equation}
 c = \operatornamewithlimits{argmin}\limits_{x \in L(a^c, b^c)} ||a^c-x||
\end{equation}

\noindent where for each point $r$ on line $L(p,q)$, $||r-p|| = ||r-q||$. Moreover, for
any three consecutive sampled points $c_i$, $c_j$ and $c_k$ on curve $C$, the arc lengths
of the curves between their closest projections are the same:
\begin{equation}
\small
 D^A(a^{c_i},a^{c_j}) + D^B(b^{c_i},b^{c_j}) = D^A(a^{c_j},a^{c_k}) + D^B(b^{c_j},b^{c_k})
\end{equation}
\noindent where $D^Z(x,y)$ is the distance along the curve $Z$ between points $x,y \in Z$.

Semantically, we want to find the curve that is composed of the loci of the smallest spheres that touch the two inputs curves A and B,
and sample it such that for each sample on the curve, the distances traveled by the matching samples along their curves is a constant.

\section{Input}
The input is the six control points for two curves $A$ and $B$, named $A_0$ to $A_5$ and $B_0$ to $B_5$, such that the first and the last control points are the same.

\section{Overview}
The project is composed of the following three main parts: (1) the representation of the curves,
(2) the computation of the average curve, and (3) the visualization of the different curve properties.
The challenge in the representation is that we want the spline curves to first meet at two end points and then,
to have $C^1$ continuity at the intersection of the splines. For the curve average, we had to ensure
the points satisfied the constraints in Equations 1-3. Lastly, the visualization contained multiple challenges
including parallel transport, circular arc computation and etc.

\section{Curve Representation}
We compose each curve out of piecewise quadratic Hermite and cubic Hermite splines. Each spline is connected at a control point by the same local velocity. This leads to a $\mathcal{C}^1$-smooth curve.
We define the tangent/velocity at control point $A_i$ as $T_i = c*(A_{i+1}-A_{i-1})$. The parameter $c$ defines the curvature or the influence of that velocity. In our experiments, we set $c$ to 0.5.

A cubic Hermite spline requires velocities at both end points, a quadratic Hermite spline only at one end point. Therefore, we use the quadratic form for the first and the last part of the curve and the cubic form for all middle parts.

\subsection{Quadratic Hermite Spline}
Given the points $X_0$ and $X_1$ and the tangent $T_0$, find a quadratic spline $P(t)$ so that the following holds:
\begin{equation}
\small
 P(0)=X_0, P(1)=X_1, P'(0)=T_0
\end{equation}
Solving this system of equation leads to the following formula:
\begin{equation}
\small
 P(t) = (t^2-2t+1)X_0 + (-t^2+2t)X_1 + (t^2-t)T_0
\label{eq:QuadraticHermite}
\end{equation}

\subsection{Cubic Hermite Spline}
Given the points $X_0$ and $X_1$ and the tangents $T_0$ and $T_1$, find a cubic spline $P(t)$ so that the following holds:
\begin{equation}
\small
 P(0)=X_0, P(1)=X_1, P'(0)=T_0, P'(1)=T_0
\end{equation}
This equation is solved similar to the quadratic case, leading to:
\begin{equation}
\small
 P(t) = (2t^3-3t^2+1)X_0 + (t^3-2t^2+t)T_0 + (-2t^3+3t^2)X_1 + (t^3-t^2)T_1
\label{eq:CubicHermite}
\end{equation}

\section{Average Curve Computation}

\subsection{Distance functions}
\subsection{Average Distance Sampling}

\section{Visualization and Editing}

\begin{figure*} %This just must appear at the page before the chapter "Displaying the average curve"
	\centering
		\begin{tabular}{cccc}
			\includegraphics[scale=0.5]{images/NoGeodesicSampling.png} & \includegraphics[scale=0.5]{images/GeodesicSampling.png} 
				& \includegraphics[scale=0.5]{images/ClosestProjections.png} & \includegraphics[scale=0.5]{images/CircularArcNet.png} \\
			a) no geodesic sampling & b) with geodesic sampling & c) the closest projections & d) net of circular arcs
		\end{tabular}
	\caption{Geodesic sampling, closest projections and circular arcs}
	\label{fig:GeodesicSampling}
\end{figure*}

We provide a framework for editing the control curve and visualizing the curve average.
More specific, we support the following features: 
\begin{itemize}
	\item Moving the control points of the two input curve
	\item Displaying the curve average with or without geodesic sampling
	\item Displaying the closest projections
	\item Displaying the circular arc and animating the curve average along it
	\item Showing the inflation tube, the envelope of the smallest spheres
\end{itemize}
We now describe the single features in more detail

\subsection{Editing and displaying the control curve}
We display the control points as gray spheres and the two control curves in green and blue.
The user can click the control points and change their positions by dragging them over the screen.

\subsection{Displaying the average curve}
The curve average is displayed by connecting the sample points using piecewise straight tubes. The faces of the tubes are rendered in alternating black and gray so you can see where the trace points are.
The user has now the option to toggle geodesic sampling on or off. The effects are shown in Fig.\ref{fig:GeodesicSampling}ab.

As you can see, without geodesic sampling, the distances between the samples of the curve average are almost equispaced, but the closest projections on the control curve, indicated by the circular arcs, vary greatly in the distances. With geodesic sampling turned on, the curve average is sampled in that way that the sum of the distances between two consecutive closest projects on the control curves are constant. However, this leads to large variation of the distances of the curve average samples.

\subsection{ClosestProjections}
To visualize the closest projections, i.e. the points on the control curve that are closest to the average curve at this point, we draw straight lines between the sample of the curve average and its closest projections. These can be seen in Fig.\ref{fig:GeodesicSampling}c. Note that the lines to both control curves are of the same length and that they stand orthogonal to the control curve at the intersections. These are the properties that define the closest projections and the curve average.

\section{Results}

\section{Future work}

\end{document}

### 2015-10-30 src/curveavg/Curve.java

@@ -11,7 +11,7 @@ package curveavg;
  */
 public class Curve {
 	public static final float EPSILON = 0.001f;
-	public static final float TANGENT_SCALE = 1f;
+	public static final float TANGENT_SCALE = 0.5f;
 	
 	//TODO: compute the curve interpolation uniform distributed over the arc-length
 	

### 2015-10-30 paper/paper.tex
added first subchapters of the visualization part to the paper

@@
 \subsection{Distance functions}
-\subsection{Average Distance Sampling}
+TODO
+
+\subsection{Curve tracing}
+TODO
+
+\subsection{Average Distance (geodesic) Sampling}
+TODO
 
@@
 
-\subsection{ClosestProjections}
+\subsection{Closest projections}
 To visualize the closest projections, i.e.
@@
 
+\subsection{Circular arc}
+Instead of displaying the closest projections as straight lines, we could also draw them using circular arcs.
+\begin{figure}[ht!] 
+  \centering
+	\begin{tikzpicture}[scale=0.5]
+		\coordinate[label=right:P] (P) at (0,0);
+		\coordinate[label=left:A] (A) at (-6,-2);
+		\coordinate[label=right:B] (B) at (6,-2);
+		\coordinate[label=right:C] (C) at (0,-20);
+		\draw (P)--(A); %closest projections
+		\draw (P)--(B);
+		\draw[blue,thick] (6.4,-0.8) coordinate -- (5.6,-3.2) coordinate; %parts of the control curve
+		\draw[green,thick] (-6.4,-0.8) coordinate -- (-5.6,-3.2) coordinate;
+		\draw (-6.2,-1.4) coordinate -- (-5.6,-1.2) coordinate -- (-5.2,-2.4) coordinate -- (-5.8,-2.6) coordinate; % right angles
+		\draw (6.2,-1.4) coordinate -- (5.6,-1.2) coordinate -- (5.2,-2.4) coordinate -- (5.8,-2.6) coordinate;
+		\path[blue,thick] (6.4,-0.8) coordinate edge[out=70,in=-120] (6.8,1) coordinate; % control curve
+		\path[blue,thick] (5.6,-3.2) coordinate edge[out=250,in=100] (5.8,-5) coordinate;
+		\path[green,thick] (-6.4,-0.8) coordinate edge[out=110,in=-120] (-7,1) coordinate;
+		\path[green,thick] (-5.6,-3.2) coordinate edge[out=290,in=80] (-5,-5) coordinate;
+		\path[black,thick] (0,0) coordinate edge[out=90,in=-70] (-0.2,1) coordinate; % average curve
+		\path[black,thick] (0,0) coordinate edge[out=-90,in=110] (0.2,-3) coordinate;
+		\draw (A)--(C); %circle sides
+		\draw (B)--(C);
+		\coordinate[label=left:r] (r) at (-4,-8);
+		\draw[red,thick] ([shift=(71.567:19)] 0,-20) arc (71.567:108.435:19); %circular arc
+		\tikzset{mypoints/.style={fill=white,draw=black,thick}} %coordinate points
+		\fill[mypoints] (P) circle (5.0pt);
+		\fill[mypoints] (A) circle (5.0pt);
+		\fill[mypoints] (B) circle (5.0pt);
+		\fill[mypoints] (C) circle (5.0pt);
+	\end{tikzpicture}
+	\caption{circular arc}
+	\label{fig:CircularArc}
+\end{figure}
+In Fig.\ref{fig:CircularArc}, you can see parts of the two input curves in green and blue and the curve average in black in the middle.
+The current trace point is $P$ and the closest projections of $P$ on the control points are $A$ and $B$.
+\\
+We can now construct a unique circle in 3d with center $C$, radius $r$ and normal $N$ that goes through $A$ and $B$ and the tangents at these points are $AP$ and $BP$.
+
+First, assume $A$,$B$,$P$ are not collinear.
+Then $ABP$ defines a plane with the normal vector $N=\underline{PA} \times \underline{PB}$.
+The circular arc lies in that plane.
+The vectors $U=N \times \underline{PA}$ and $V=N \times \underline{PB}$ are normals of the circle at these points.
+Therefore, the rays $A+\alpha U$ and $B+\beta V$ intersect exactly at the circle center $C$.
+This system is overdetermined but we can solve for $\alpha$ in the following way:
+\begin{equation}
+\begin{array}{rcl}
+A + \alpha U &=& B + \beta V \\
+\alpha U &=& B-A+\beta V \\
+(\alpha U) \times V &=& (AB+\beta V) \times V \\
+\alpha (U\times V) &=& (AB) \times V
+\end{array}
+\label{eq:CircArc}
+\end{equation}
+And now we have $\alpha = \frac{ ((AB)\times V).x }{ (U \times V).x}$. We can pick any coordinate, all have to be the same.
+
+This then leads to $C=A + \alpha U$, $r=|AC|$ and $N$ as described above.
+
+To draw the circular arc by approximating it using piecewise straight tubes.
+We sample the arc by rotating the vector $CA$ around $N$ until it becomes $CB$.
+This is done the same way as in the Swirl-project:
+\begin{equation}
+\begin{array}{lcl}
+ CA^{\circ}(\theta; N) &:=& C + U^{\circ}(\alpha; N) \\
+                        & =& C + \cos \alpha U - \sin \alpha (N \times U)\\
+ \multicolumn{3}{l}{\text{with } V := CA\angle\underline{N} = (W\bullet\underline{N})\underline{N}} \\
+ \multicolumn{3}{l}{\text{and } U = CA-V} \\
+ \multicolumn{3}{l}{\theta \text{ ranges from } 0 \text{ to } CA^{\wedge}CB=\cos^{-1}(\underline{CA}\bullet\underline{CB})}
+\end{array}
+\label{eq:Rot}
+\end{equation}
+
+Furthermore, if we sample the circular arcs at some angles and combine the samples with the same angle, we obtain a net, as shown in Fig.\ref{GeodesicSampling}d.
+If both control curves lie in a plane, this net also lies in the same plane.
+
+We further provide the ability to animate the average curve between the two control curves.
+In that animation, each sample of the average curve follows the circular arc between the closest projections.
+
+\subsection{Inflation tube}
+TODO
+
 \section{Results}

### 2015-10-30 src/curveavg/CurveAverage.java

@@ -130,7 +130,7 @@ public class CurveAverage extends AbstractPApplet {
 		recalculateCurve = true;
 		
 		//test
-		spiralControlPoints();
+//		spiralControlPoints();
 
 		rx += 0.0001f;
 		ry += 0.0001f;

### 2015-10-30 src/curveavg/CurveAverage.java
added four presets

@@ -81,7 +81,7 @@ public class CurveAverage extends AbstractPApplet {
 
 	String title = "6491 P3 2015: Curve Average", name = "Sebastian Weiß, Can Erdogan",
 			menu = "!:picture, ~:(start/stop)capture, space:rotate, "
-			+ "s/wheel:closer, a:anim, x:show coordinate axes\n"
+			+ "s/wheel:closer, a:anim, x:show coordinate axes, 1-4:load presets\n"
 			//+ "i:interpolate control curve, e:equispaced interpolation, m:show medial axis, "
 			+ "e:equispaced interpolation, m:show medial axis, g:geodesic sampling of the MA, "
 			+ "p:show closest projections, c:show circular arc, n:show net (implies c)\n"
@@ -103,19 +103,12 @@ public class CurveAverage extends AbstractPApplet {
 		myFace = loadImage("data/pic.jpg");  // load image from file pic.jpg in folder data *** replace that file with your pic of your own face
 		textureMode(NORMAL);
 
-		//example curve
-		controlPoints = new Vector3f[]{
-			new Vector3f(-60, -40, -80), //start
-			new Vector3f(-60, -10, -40),
-			new Vector3f(-30, 12, -6),
-			new Vector3f(0, 50, 60),
-			new Vector3f(80, 45, 140),
-			new Vector3f(170, 90, 200), //end
-			new Vector3f(140, 70, 160),
-			new Vector3f(100, 37, 100),
-			new Vector3f(50, 19, 40),
-			new Vector3f(0, 15, -20)
-		};
+		controlPoints = new Vector3f[10];
+		for (int i=0; i<10; ++i) {
+			controlPoints[i] = new Vector3f();
+		}
+		//first setup
+		setup1_Normal();
 //		for(int i = 0; i < controlPoints.length; i++) controlPoints[i].y = 0.0f;
 
 		curveA = new Vector3f[controlPoints.length / 2 + 1];
@@ -148,7 +141,19 @@ public class CurveAverage extends AbstractPApplet {
 		 };
 		 */
 	}
-	private void spiralControlPoints() {
+	private void setup1_Normal() {
+		controlPoints[0].set(-60, -40, -80);
+		controlPoints[1].set(-60, -10, -40);
+		controlPoints[2].set(-30, 12, -6);
+		controlPoints[3].set(0, 50, 60);
+		controlPoints[4].set(80, 45, 140);
[…]

### 2015-10-30 paper/paper.tex
started with the inflation chapter

@@
 		\begin{tabular}{cccc}
-			\includegraphics[scale=0.5]{images/NoGeodesicSampling.png} & \includegraphics[scale=0.5]{images/GeodesicSampling.png} 
-				& \includegraphics[scale=0.5]{images/ClosestProjections.png} & \includegraphics[scale=0.5]{images/CircularArcNet.png} \\
+			\includegraphics[scale=0.4]{images/NoGeodesicSampling.png} & \includegraphics[scale=0.4]{images/GeodesicSampling.png} 
+				& \includegraphics[scale=0.4]{images/ClosestProjections.png} & \includegraphics[scale=0.4]{images/CircularArcNet.png} \\
+			\small
 			a) no geodesic sampling & b) with geodesic sampling & c) the closest projections & d) net of circular arcs
@@
 
-To draw the circular arc by approximating it using piecewise straight tubes.
-We sample the arc by rotating the vector $CA$ around $N$ until it becomes $CB$.
+If the points $A$,$B$,$P$ are collinear, indicated by $N=0$ (plus epsilon), then the circular arc degenerates to a linear interpolation (lerp) from A to B through P.
+
+We draw the circular arc by approximating it using piecewise straight tubes.
+The samples are obtained by rotating the vector $CA$ around $N$ until it becomes $CB$.
 This is done the same way as in the Swirl-project:
@@
 \subsection{Inflation tube}
-TODO
+\begin{figure}
+	\centering
+		\begin{tabular}{ccc}
+			\includegraphics[scale=0.7]{images/InflationNoParallelTransport.png} & \includegraphics[scale=0.7]{images/InflationParallelTransport.png} 
+				& \includegraphics[scale=0.7]{images/InflationHalf.png}\\
+			a) no  & b) with & c) \\
+			parallel transport & parallel transport & half tube
+		\end{tabular}
+	\caption{the inflation tube}
+	\label{fig:Inflation}
+\end{figure}
+As a last graphic feature, we draw the inflation tube. The inflation tube is the envelope of the minimal balls along the curve average.
+It can be seen as the union of all balls with the center on the curve average that exactly touches the control curve and contains no other point in the interior.
+
+TODO: displaying: NO piecewise tubes orthogonal to the curve average, but rotating PA around the curve average (image)
+TODO: parallel transport
 

### 2015-10-30 src/curveavg/CurveAverage.java

@@ -56,6 +56,7 @@ public class CurveAverage extends AbstractPApplet {
 	private boolean showNet = false;
 	private boolean showInflation = false;
 	private boolean showInflationWireframed = false;
+	private boolean inflationParallelTransport = true;
 	private boolean showHalfInflation = false;
 	private int animDirection;
 	private float animStep;
@@ -85,7 +86,7 @@ public class CurveAverage extends AbstractPApplet {
 			//+ "i:interpolate control curve, e:equispaced interpolation, m:show medial axis, "
 			+ "e:equispaced interpolation, m:show medial axis, g:geodesic sampling of the MA, "
 			+ "p:show closest projections, c:show circular arc, n:show net (implies c)\n"
-			+ "t:show inflation tube, w:wireframed inflation tube, h:show only half of the inflation tube",
+			+ "t:show inflation tube, w:wireframed inflation tube, l:enable parallel transport, h:show only half of the inflation tube",
 			guide = "click'n'drag the control points of the two curves green and blue"; // user's guide
 
 	public static void main(String[] args) {
@@ -511,7 +512,7 @@ public class CurveAverage extends AbstractPApplet {
 			}
 			showInflation(A, B, medialAxis, radii, 
 					showHalfInflation ? MEDIAL_AXIS_INFLATION_RESOLUTION/2 : MEDIAL_AXIS_INFLATION_RESOLUTION, 
-					grey80, lightgrey80, showInflationWireframed, showHalfInflation, !showHalfInflation);
+					grey80, lightgrey80, showInflationWireframed, showHalfInflation, !showHalfInflation && inflationParallelTransport);
 		}
             
 	}
@@ -842,6 +843,9 @@ public class CurveAverage extends AbstractPApplet {
 		}
 		if (key == 'w') {
 			showInflationWireframed = !showInflationWireframed;
+		}
+		if (key == 'l') {
+			inflationParallelTransport = !inflationParallelTransport;
 		}
 		if (key == 'h') {
 			showHalfInflation = !showHalfInflation;

### 2015-10-30 paper/paper.tex
added inflation tube chapter

@@
 
-TODO: displaying: NO piecewise tubes orthogonal to the curve average, but rotating PA around the curve average (image)
-TODO: parallel transport
+\subsubsection{Drawing the inflation tube}
+
+To draw the inflation tube, we get the following input from the tracing algorithm: the samples of the average curve $P_i$, $i\in\{0,...n-1\}$, the closest projections on the two control curves $A_i$ and $B_i$ and the radii or distances to the closest projections $r_i$.
+We further define the tangent $T_i$ at the average curve sample $i$ as $T_i=\frac{1}{2}(P_{i+1}-P_{i-1})$.
+
+The first way one might think of drawing it, is to circles around the average curve samples $P_i$ in the plane orthogonal to the tangents $T_i$ with radii $r_i$ and connect them to tubes.
+However, this approach will lead to wrong results since the closest projections to the control curves are not orthogonal to the average curve tangent.
+Therefore, the distance from the closest projections to the average curve is not equal to the radius of the circle when it should touch the control curves.\\
+
+We present a different approach now.
+Instead of drawing circles orthogonal to the tangents, we directly take the vectors $P_i A_i$ and rotate them around $T_i$.
+This technique is visualized in Fig.\ref{fig:InflationDrawing}.
+We rotate $P_i A_i$ as described in Eq.\ref{eq:Rot} the whole $360^{\circ}$.
+At $180^{\circ}$, the vector becomes $P_i B_i$.
+We then sample the circles at constant angles and connect samples with the same angle, leading to the tube as shown in Fig.\ref{fig:Inflation}a.
+
+\begin{figure}[ht!] 
+  \centering
+	\begin{tikzpicture}[scale=0.5]
+		\draw[black,thick] (0,0) coordinate -- (0,10) coordinate; %curves
+		\draw[green,thick] (-3,0) coordinate -- (-6,10) coordinate;
+		\draw[blue,thick] (3,0) coordinate -- (6,10) coordinate;
+		\coordinate[label=right:$P_{i+1}$] (P2) at (0,8) coordinate; %samples
+		\coordinate[label=right:$P_{i}$] (P1) at (0,6) coordinate;
+		\coordinate[label=right:$P_{i-1}$] (P0) at (0,4) coordinate;
+		\draw[dashed] (0,8) coordinate -- (5,6.5) coordinate; %closest projections
+		\draw[dashed] (0,8) coordinate -- (-2.5,7.25) coordinate node[below] {$r_{i+1}$} -- (-5,6.5) coordinate;
+		\draw[dashed] (0,6) coordinate -- (4.5,4.65) coordinate;
+		\draw[dashed] (0,6) coordinate -- (-2.25,5.325) coordinate node[below] {$r_{i}$} -- (-4.5,4.65) coordinate;
+		\draw[dashed] (0,4) coordinate -- (4,2.8) coordinate;
+		\draw[dashed] (0,4) coordinate -- (-2,3.4) coordinate node[below] {$r_{i-1}$} -- (-4,2.8) coordinate;
+		\coordinate[label=left:$A_{i+1}$] (A2) at (-5,6.5) coordinate; %closest projection points
+		\coordinate[label=left:$A_{i}$] (A1) at (-4.45,4.65) coordinate;
+		\coordinate[label=left:$A_{i-1}$] (A0) at (-3.9,2.8) coordinate;
+		\coordinate[label=left:$B_{i+1}$] (B2) at (5,6.5) coordinate;
+		\coordinate[label=left:$B_{i}$] (B1) at (4.45,4.65) coordinate;
+		\coordinate[label=left:$B_{i-1}$] (B0) at (3.9,2.8) coordinate;
+		\draw[red,thick] ([shift=(180:5)] 0,6.5) arc (180:0:5 and 3); %tube elipses
+		\draw[red,dashed,thick] ([shift=(180:5)] 0,6.5) arc (180:360:5 and 3);
+		\draw[red,thick] ([shift=(180:4.5)] 0,4.65) arc (180:0:4.5 and 2.7);
+		\draw[red,dashed,thick] ([shift=(180:4.5)] 0,4.65) arc (180:360:4.5 and 2.7);
+		\draw[red,thick] ([shift=(180:3.9)] 0,2.8) arc (180:0:3.9 and 2.4);
+		\draw[red,dashed,thick] ([shift=(180:3.9)] 0,2.8) arc (180:360:3.9 and 2.4);
+		\tikzset{mypoints/.style={fill=white,draw=black,thick}} %coordinate points
+		\foreach \p in {P0,P1,P2,A0,A1,A2,B0,B1,B2}
+			\fill[mypoints] (\p) circle (5.0pt);
+	\end{tikzpicture}
+	\caption{inflation tube}
+	\label{fig:InflationDrawing}
+\end{figure}
+
+\subsubsection{Parallel transport}
+When the control curves rotate around the average curve, the inflation tube follows this rotation, as seen in Fig.\ref{fig:Inflation}a.
+However, often Fig.\ref{fig:Inflation}b is required.
+Here, the rotation of the inflation tube stays constant although the control curves rotate around it.
+
+The technique to achieve this result is called Parallel Transport.
+The only thing we have to do is to modify the start angle $\alpha_i$.
+Instead of sampling each circle (see previous section) starting with $0^{\circ}$, we start it at an angle that minimizes the rotation to the previous circle.
+
+The task is now the following: given the previous tangent $\underline{T_{i-1}}$, rotated vector $P_{i-1}A_{i-1}$ and start angle $\alpha_{i-1}$, and the current tangent $\underline{T_i}$ and rotated vector $P_i A_i$, find the current start angle $\alpha_i$ so that the twist is minimized.
+In other words, we have to find $\alpha_i$ so that $P_i A_i^{\circ}(\alpha_i; T_i)$ lies in the plane spanned by $T_{i-1}$ and $X:=P_{i-1}A_{i-1}^{\circ}(\alpha_{i-1}; T_{i-1})$:
+
+\begin{equation}
+	\begin{array}{rcl}
+		\underline{X} &:=& \underline{P_{i-1}A_{i-1}^{\circ}(\alpha_{i-1}; T_{i-1})} \\
+		\underline{N'} &:=& \underline{T_{i-1}} \times \underline{X} \\
+		\underline{PA'} &:=& \underline{P_i A_i} \\
+		x &:=& \underline{PA'} \bullet \underline{N'} \\
+		y &:=& \underline{PA'} \bullet \underline{X} \\
+		PA'' &:=& xN' + yX \\
+		\alpha_i &:=& ( (x<0)?(-1):1 ) \cdot \cos^{-1}(\underline{X} \bullet \underline{PA''})
+	\end{array}
+\label{eq:ParallelTransport}
+\end{equation}
+$\underline{N'}$ is the normal of the reference plane, then we project $P_i A_i$ in that plane and obtain $PA''$.
+The final angle $\alpha_i$ is then the angle between the reference vector $X$ and the projected vector $PA''$.
+Since this angle does not include the direction of the twist, we have to invert the angle if $x<0$.
+
+\subsubsection{half tube}
+Instead of drawing the whole inflation tube without parallel transport, we could also draw only half of it. This leads to the interesting effect of a waterslide as seen in Fig.\ref{fig:Inflation}c.
 

### 2015-10-30 src/curveavg/CurveAverage.java

@@ -705,7 +705,7 @@ public class CurveAverage extends AbstractPApplet {
 					Vector3f reference = rotate(angles[i-1], CAp, Ns[i-1]);
 					//compute the angle so that CA°angle lies in the plane N x reference
 					Vector3f refPrime = reference.normalize();
-					Vector3f NPrime = N.cross(refPrime);
+					Vector3f NPrime = Ns[i-1].cross(refPrime);
 					Vector3f CAPrime = CA.normalize();
 					float x = CAPrime.dot(NPrime);
 					float y = CAPrime.dot(refPrime);

### 2015-11-01 src/curveavg/MedialAxisTransform.java
Changed the Laguerre solvers discretization from 0.001 to 0.01 and timing is much better

@@ -23,7 +23,7 @@ import org.apache.commons.math3.exception.TooManyEvaluationsException;
  */
 public class MedialAxisTransform {
 	private static final Logger LOG = Logger.getLogger(MedialAxisTransform.class.getName());
-        private static final float STEPSIZE = 2f;
+        private static final float STEPSIZE = 1f;
         private static final int MAX_MA_ITERATION = 500;
         
 	public static class Line {
@@ -615,14 +615,14 @@ public class MedialAxisTransform {
 		SolverResult res = solveQuinticLaguerre(a / f, b / f, c / f, d / f, e / f, 1.0f);
 
 		// Check which endpoint is closer
-		float dist0 = Q.distance(P0);
-		float dist1 = Q.distance(P1);
-		float minEndpointDist = Math.min(dist0, dist1);
+//		float dist0 = Q.distance(P0);
+//		float dist1 = Q.distance(P1);
+//		float minEndpointDist = Math.min(dist0, dist1);
 //            System.out.println("Minimum endpoint dist: " + minEndpointDist);
 
             // For each nonimaginary root, with real part within [0,1], compute 
 		// the distance to the input point and find the smallest value
-		float bestDist = 1e6f;
+		float bestDistSQ = 1e6f;
 		float time = -1f;
 		for (int i = 0; i < 6; i++) {
 			if (Math.abs(res.im[i]) > 1e-3) {
@@ -632,11 +632,11 @@ public class MedialAxisTransform {
 				continue;
 			}
 			Vector3f Pt = Curve.cubicHermite(P0, T0, P1, T1, res.re[i]);
-			float dist = Q.distance(Pt);
+			float distSQ = Q.distanceSquared(Pt);
 //                System.out.println("\troot t: " + res.re[i] + ", pt: " + Pt.toString() + ", dist:" + dist);
-			if (dist < bestDist) {
+			if (distSQ < bestDistSQ) {
 				time = res.re[i];
-				bestDist = dist;
+				bestDistSQ = distSQ;
 			}
 		}
 
@@ -668,7 +668,7 @@ public class MedialAxisTransform {
 		double minBound = 0.0;
 		double sol;
 		int root_idx = 0;
-		final double dt = 0.001;
+		final double dt = 0.01;
 		for (double t = -dt; t < 1.0; t += dt) {
 
 			// Check if there is a sign-crossing between the two times

### 2015-11-01 src/curveavg/Curve.java
Got the incompatible coloring - kind of butchered the interpolate function definition

@@ -5,6 +5,11 @@
  */
 package curveavg;
 
+import java.util.ArrayList;
+import java.util.List;
+import processing.core.*;
+import processing.event.*;
+
 /**
  * Implements the curve interpolation using piecewise Hermite.
  * @author Sebastian Weiss
@@ -102,7 +107,9 @@ public class Curve {
 	 * @return equispaced samplings of the interpolated curve.
 	 * @see #interpolate(curveavg.Vector3f[], float) 
 	 */
-	public static Vector3f[] interpolateEquispacedArcLength(Vector3f[] controlPoints, int numSamples) {
+	public static Vector3f[] interpolateEquispacedArcLength(Vector3f[] controlPoints, int numSamples,
+                List <Vector2f> incompatibleSections, List <Integer> colors, int defaultColor) {
+            
 		//First step: Oversample curve to calculate the total arc length
 		Vector3f[] samples1 = new Vector3f[numSamples * 2];
 		float step1 = samples1.length-1;
@@ -121,6 +128,7 @@ public class Curve {
 		//Third step: sample final curve equispaced
 		Vector3f[] samples2 = new Vector3f[numSamples];
 		samples2[0] = controlPoints[0]; //fix first and last point
+                colors.add(AbstractPApplet.blue);
 		samples2[numSamples-1] = controlPoints[controlPoints.length-1];
 		int j=0;
 		float step2 = arcLength / (numSamples-1);
@@ -135,7 +143,21 @@ public class Curve {
 			}
 			float t = (j + (d-distances[j])/(distances[j+1]-distances[j])) / step1;
 			samples2[i] = interpolate(controlPoints, t);
+                        
+                        // Change the color if in bad section
+                        boolean inBadSection = false;
+                        for(int sec_idx = 0; sec_idx < incompatibleSections.size(); sec_idx++) {
+                            if(t >= incompatibleSections.get(sec_idx).x && t <= incompatibleSections.get(sec_idx).y) {
+                                inBadSection = true;
+                                break;
+                            }
+                        }
+                        if(inBadSection)
+                            colors.add(AbstractPApplet.red);
+                        else 
+                            colors.add(defaultColor);
 		}
+                colors.add(AbstractPApplet.blue);
 		return samples2;
 	}
 	

### 2015-11-01 src/curveavg/CurveAverage.java

@@ -50,7 +50,7 @@ public class CurveAverage extends AbstractPApplet {
 	private boolean interpolateControlCurve = true;
 	private boolean equispacedInterpolation = true;
 	private boolean showMedialAxis = true;
-	private boolean geodesicMedialAxis = false;
+	private boolean geodesicMedialAxis = true;
 	private boolean showClosestProjection = false;
 	private boolean showCircularArcs = true;
 	private boolean showNet = false;
@@ -78,8 +78,12 @@ public class CurveAverage extends AbstractPApplet {
 	public Vector3f[] debugPoints;             ///< Some useful points that we can drag around 
 	boolean recalculateCurve = false;
 	private Vector3f[] samplesA, samplesB;
+        List <Integer> colorsA = new ArrayList<Integer>();
+        List <Integer> colorsB = new ArrayList<Integer>();
 	List<MedialAxisTransform.TracePoint> ma = new ArrayList<MedialAxisTransform.TracePoint>();
-
+        List<Vector2f> incompatibleSectionsA = new ArrayList<Vector2f>();
+        List<Vector2f> incompatibleSectionsB = new ArrayList<Vector2f>();
+        
 	String title = "6491 P3 2015: Curve Average", name = "Sebastian Weiß, Can Erdogan",
 			menu = "!:picture, ~:(start/stop)capture, space:rotate, "
 			+ "s/wheel:closer, a:anim, x:show coordinate axes, 1-4:load presets\n"
@@ -347,8 +351,10 @@ public class CurveAverage extends AbstractPApplet {
 			//recalculate control curves
 			if (interpolateControlCurve) {
 				if (equispacedInterpolation) {
-					samplesA = Curve.interpolateEquispacedArcLength(curveA, (curveA.length - 1) * CURVE_INTERPOLATION_SAMPLES);
-					samplesB = Curve.interpolateEquispacedArcLength(curveB, (curveB.length - 1) * CURVE_INTERPOLATION_SAMPLES);
+                                        colorsA.clear();
+                                        colorsB.clear();
[…]

### 2015-11-01 src/curveavg/MedialAxisTransform.java

@@ -148,9 +148,11 @@ public class MedialAxisTransform {
 	 * @param curveB the control points of the second curve
 	 * @param resolution the expected resolution of the target curve
 	 * @param output an empty list, place the traced point in here
+         * @param incompatibleSections a list of Vector2f for the beginning and 
+         * end times of the bad sections
 	 */
 	public static void geodesicTrace(Vector3f[] curveA, Vector3f[] curveB,
-			int resolution, List<TracePoint> output) {
+			int resolution, List<TracePoint> output, List <Vector2f> incompatibleSectionsA, List <Vector2f> incompatibleSectionsB) {
 		//trace it normally
 		List<TracePoint> traceList = new ArrayList<>();
 		trace(curveA, curveB, traceList);
@@ -211,7 +213,14 @@ public class MedialAxisTransform {
 			float realDist = 
 					Curve.computeArcLength(curveA, start.projectionOnA[0], r.ta, 32)
 					+ Curve.computeArcLength(curveB, start.projectionOnB[0], r.tb, 32);
-			System.out.println("real step size: "+realDist);
+                        if(realDist < 1e-3)
+                            System.out.println("real step size: "+realDist);
+                        
+                        if(realDist > targetStepSize * 1.5) {
+                            incompatibleSectionsA.add(new Vector2f(start.projectionOnA[0], r.ta));
+                            incompatibleSectionsB.add(new Vector2f(start.projectionOnB[0], r.tb));
+                        }
+                        
 			//add add trace point
 			TracePoint tp = new TracePoint(r.center, r.radius, r.ta, r.tb);
 			output.add(tp);
@@ -397,13 +406,13 @@ public class MedialAxisTransform {
 			current = r.center;
 //			System.out.println("current="+current+", tA="+nextTA+", tB="+nextTB);
 			if (nextTA >= 1 || nextTB >= 1) {
-                                LOG.info("reached end");
+//                                LOG.info("reached end");
 				return true; //end reached
 			}
 			if (nextTA < ta || nextTB < tb) {
 				//use this as a stop condition
 				
-				LOG.severe("going backwards!!!");
+//				LOG.severe("going backwards!!!");
                                 return true;
 //				return false;
 			}

### 2015-11-01 src/curveavg/CurveAverage.java
some stability improvements

@@ -217,8 +217,8 @@ public class CurveAverage extends AbstractPApplet {
 		controlPoints[4].set(0, -20, -70);
 		controlPoints[5].set(150, -70, -100);
 		controlPoints[6].set(-16.6f, -56.2f, -56.8f);
-		controlPoints[7].set(-73.5f, 57.4f, -20.8f);
-		controlPoints[8].set(46.1f, 28, -7.4f);
+		controlPoints[7].set(-53.5f, 47.4f, -20.8f);
+		controlPoints[8].set(30.1f, 28, -7.4f);
 		controlPoints[9].set(-30.4f, -55f, 40);
 	}
 
@@ -422,9 +422,13 @@ public class CurveAverage extends AbstractPApplet {
 			}
 		}
         
+		if (ma.size() < 10) {
+			return; //unable to trace medial axis
+		}
+		
                 //show circular arcs
 		if (showCircularArcs || showNet) {
-                        assert(ma.size() >= MEDIAL_AXIS_ARC_COUNT);     // to ensure the division below makes sense and step is not 0
+            //assert(ma.size() >= MEDIAL_AXIS_ARC_COUNT);     // to ensure the division below makes sense and step is not 0
 			int step = (int) (ma.size() / (float) MEDIAL_AXIS_ARC_COUNT);
             
 			for (float f=step; f<ma.size(); f+=step) {

### 2015-11-01 src/curveavg/MedialAxisTransform.java

@@ -213,8 +213,10 @@ public class MedialAxisTransform {
 			float realDist = 
 					Curve.computeArcLength(curveA, start.projectionOnA[0], r.ta, 32)
 					+ Curve.computeArcLength(curveB, start.projectionOnB[0], r.tb, 32);
-                        if(realDist < 1e-3)
-                            System.out.println("real step size: "+realDist);
+                        if(realDist < 1e-3) {
+							LOG.warning("zero length step");
+							i++;
+						}
                         
                         if(realDist > targetStepSize * 1.5) {
                             incompatibleSectionsA.add(new Vector2f(start.projectionOnA[0], r.ta));

### 2015-11-01 src/curveavg/CurveAverage.java
some more stability improvements, no endless loops should happen now

@@ -392,6 +392,15 @@ public class CurveAverage extends AbstractPApplet {
 	}
 	
 	private void showMedialAxis() {
+		//test if it was possible to trace the medial axis
+		if (ma.size() < 10) {
+			return; //unable to trace medial axis
+		}
+		for (MedialAxisTransform.TracePoint p : ma) {
+			if (Float.isNaN(p.center.x) || Float.isNaN(p.center.y) || Float.isNaN(p.center.z)) {
+				return;
+			}
+		}
             
 		//interpolate center polyline
 		if (showMedialAxis) {
@@ -421,15 +430,12 @@ public class CurveAverage extends AbstractPApplet {
 				}
 			}
 		}
-        
-		if (ma.size() < 10) {
-			return; //unable to trace medial axis
-		}
 		
-                //show circular arcs
+		//show circular arcs
 		if (showCircularArcs || showNet) {
             //assert(ma.size() >= MEDIAL_AXIS_ARC_COUNT);     // to ensure the division below makes sense and step is not 0
 			int step = (int) (ma.size() / (float) MEDIAL_AXIS_ARC_COUNT);
+			step = Math.max(1, step);
             
 			for (float f=step; f<ma.size(); f+=step) {
 
@@ -644,22 +650,20 @@ public class CurveAverage extends AbstractPApplet {
 				P[p][j].addScaledLocal(s[j], J);
 			}
 			p = 1 - p;
-			if (i > 0) {
-				for (int j = 0; j < ne; j++) {
-					if (dark && checked) {
-						fill(200, 200, 200);
-					} else {
-						fill(col);
-					}
-					dark = !dark;
-					int jp = (j + ne - 1) % ne;
-					beginShape(QUADS);
-					vertex(P[p][jp].mult(0.5f));
-					vertex(P[p][j].mult(0.5f));
-					vertex(P[1 - p][j].mult(0.5f));
-					vertex(P[1 - p][jp].mult(0.5f));
-					endShape(CLOSE);
+			for (int j = 0; j < ne; j++) {
+				if (dark && checked) {
+					fill(200, 200, 200);
+				} else {
+					fill(col);
 				}
+				dark = !dark;
+				int jp = (j + ne - 1) % ne;
+				beginShape(QUADS);
+				vertex(P[p][jp].mult(0.5f));
+				vertex(P[p][j].mult(0.5f));
+				vertex(P[1 - p][j].mult(0.5f));
[…]

### 2015-11-01 src/curveavg/MedialAxisTransform.java

@@ -170,7 +170,16 @@ public class MedialAxisTransform {
 		output.clear();
 		output.add(trace[0]);
 		int i=0;
+		int maxIterations = 1000;
+		int iterationCounter = 0;
 		while (true) {
+			//in rare cases, this produces endless loops
+			if (iterationCounter > maxIterations) {
+				LOG.warning("geodesic trace took too long");
+				break;
+			}
+			iterationCounter++;
+			
 			TracePoint start = output.get(output.size()-1);
 			//find trace point that is behind the target step size
 			int lowerIndex = i;

### 2015-11-01 paper/paper.tex
added short result chapter and example images

@@
 \section{Results}
+In conclusion, we are able to trace the average curve of two piecewise hermite interpolated curves in real-time, whereat the user defines the control points by dragging them around.
+We presented ideas on how to detect if the two control curves are compatible or not.
+Furthermore, we showed different ways on how to display the average curve, using line segments to the closest projections, circular arcs or the inflation tube.
 
 \section{Future work}
+This work can easily be extended to any kind of input curve and any number of control points.
+Also more freedom for the user to specify the input curve, like editing the velocities, might be helpful for specific applications.
+
+% EXAMPLES
+\begin{figure*}
+	\centering
+		\includegraphics[scale=0.4]{images/Result1.png}
+		\includegraphics[scale=0.4]{images/Result2.png}
+	\caption{Example 1: a bow}
+	\label{fig:Result1}
+\end{figure*}
+\begin{figure*}
+	\centering
+		\includegraphics[scale=0.4]{images/Result3.png}
+	\caption{Example 2: a waterslide}
+	\label{fig:Result3}
+\end{figure*}
+\begin{figure*}
+	\centering
+		\includegraphics[scale=0.5]{images/Result4.png}
+		\includegraphics[scale=0.5]{images/Result5.png}
+	\caption{Example 3: extreme cases and non-compatibility detection}
+	\label{fig:Result4}
+\end{figure*}
 

### 2015-11-01 paper/paper.tex
proofread introduction chapter and visualization chapter

@@
 \section{Objective}
-Formally, for two curves $A = \{a_0 ... a_{n-1}\}$ and $B = \{b_0 ... b_{n-1}\}$, each with $n$ points and $\forall x \in (A \cup B), x \in \mathcal{R}^3$,
+Formally, for two curves $A$ and $B$ defined by the control points $\{A_0 ... A_{n-1}\}$ and $\{B_0 ... B_{n-1}\}$, all points are in $\mathbb{R}^3$, and an interpolation scheme,
 the goal is to find a curve $C$ such that for each point $c \in C$, and its closest projections
@@
 
-\noindent where for each point $r$ on line $L(p,q)$, $||r-p|| = ||r-q||$. Moreover, for
-any three consecutive sampled points $c_i$, $c_j$ and $c_k$ on curve $C$, the arc lengths
+\noindent where for each point $r$ on line $L(p,q)$, $||r-p|| = ||r-q||$. We call the curves $A$ and $B$ the control curves and $C$ the average curve.
+Moreover, for any three consecutive sampled points $c_i$, $c_j$ and $c_k$ on curve $C$, the arc lengths
 of the curves between their closest projections are the same:
@@
 
+It can also be seen as a generalization of the medial axis transformation from the 2D case to the 3D case: 
+In every projection of the $\mathbb{R}^3$ in the $\mathbb{R}^2$, the average curve $C$ is the medial axis of $A$ and $B$.
+
 \section{Input}
-The input is the six control points for two curves $A$ and $B$, named $A_0$ to $A_5$ and $B_0$ to $B_5$, such that the first and the last control points are the same.
+For simplicity, we only focus on a special case that each control curve is defined by six control points $A_0$ to $A_5$ and $B_0$ to $B_5$, such that the first and the last control points are the same.
 
 \section{Overview}
-The project is composed of the following three main parts: (1) the representation of the curves,
-(2) the computation of the average curve, and (3) the visualization of the different curve properties.
+The project is composed of the following three main parts: (1) the representation of the curves (see \ref{chapter:CurveRepr}),
+(2) the computation of the average curve (see \ref{chapter:Tracing}), and (3) the visualization of the different curve properties (see \ref{chapter:Visual}).
 The challenge in the representation is that we want the spline curves to first meet at two end points and then,
@@
 
-\section{Curve Representation}
+\section{Curve Representation}\label{chapter:CurveRepr}
 We compose each curve out of piecewise quadratic Hermite and cubic Hermite splines.
@@
 This leads to a $\mathcal{C}^1$-smooth curve.
-We define the tangent/velocity at control point $A_i$ as $T_i = c*(A_{i+1}-A_{i-1})$. The parameter $c$ defines the curvature or the influence of that velocity. In our experiments, we set $c$ to 0.5.
+We define the tangent/velocity at control point $A_i$ as $T_i = c(A_{i+1}-A_{i-1})$. The parameter $c$ defines the curvature or the influence of that velocity. In our experiments, we set $c$ to 0.5.
 
@@
 \begin{equation}
-\small
  P(0)=X_0, P(1)=X_1, P'(0)=T_0
@@
 \begin{equation}
-\small
  P(t) = (t^2-2t+1)X_0 + (-t^2+2t)X_1 + (t^2-t)T_0
@@
 \begin{equation}
-\small
  P(0)=X_0, P(1)=X_1, P'(0)=T_0, P'(1)=T_0
@@
 \begin{equation}
-\small
- P(t) = (2t^3-3t^2+1)X_0 + (t^3-2t^2+t)T_0 + (-2t^3+3t^2)X_1 + (t^3-t^2)T_1
+\begin{array}{rcl}
+	P(t) &=& (2t^3-3t^2+1)X_0 + (t^3-2t^2+t)T_0 \\
+	     &+& (-2t^3+3t^2)X_1 + (t^3-t^2)T_1
+\end{array}
 \label{eq:CubicHermite}
@@
 
-\section{Average Curve Computation}
+\section{Average Curve Computation}\label{chapter:Tracing}
 
@@
 
-\section{Visualization and Editing}
+\subsection{Detect non-compatible control curves}
+TODO
+
+\section{Visualization and Editing}\label{chapter:Visual}
 
@@
 As you can see, without geodesic sampling, the distances between the samples of the curve average are almost equispaced, but the closest projections on the control curve, indicated by the circular arcs, vary greatly in the distances.
-With geodesic sampling turned on, the curve average is sampled in that way that the sum of the distances between two consecutive closest projects on the control curves are constant.
+With geodesic sampling turned on, the curve average is sampled in that way that the sum of the distances between two consecutive closest projects on the control curves is constant.
 However, this leads to large variation of the distances of the curve average samples.
@@
 \\
-We can now construct a unique circle in 3d with center $C$, radius $r$ and normal $N$ that goes through $A$ and $B$ and the tangents at these points are $AP$ and $BP$.
+We can now construct a unique circle in 3d with center $C$ and radius $r$ that goes through $A$ and $B$, the tangents at these points are $AP$ and $BP$ and lies in the convex hull of $A$,$B$ and $P$.
 
@@
 \begin{array}{lcl}
- CA^{\circ}(\theta; N) &:=& C + U^{\circ}(\alpha; N) \\
-                        & =& C + \cos \alpha U - \sin \alpha (N \times U)\\
+ CA^{\circ}(\theta; N) &:=& C + U^{\circ}(\theta; N) \\
+                        & =& C + \cos \theta U - \sin \theta (N \times U)\\
  \multicolumn{3}{l}{\text{with } V := CA\angle\underline{N} = (W\bullet\underline{N})\underline{N}} \\
@@
 
-Furthermore, if we sample the circular arcs at some angles and combine the samples with the same angle, we obtain a net, as shown in Fig.\ref{GeodesicSampling}d.
+Furthermore, if we sample the circular arcs at some angles and combine the samples with the same angle, we obtain a net, as shown in Fig.\ref{fig:GeodesicSampling}d.
 If both control curves lie in a plane, this net also lies in the same plane.

### 2015-11-02 paper/paper.tex
Wrote the algorithms and the block diagram. Next: Fill in the text

@@
 
+\usepackage{blindtext}
 \usepackage[english]{babel}
@@
 \usepackage{subcaption}
-\usepackage{algorithm}
-\usepackage{algorithmic}
+% \usepackage{algorithm}
+% \usepackage{algorithmic}
 \usepackage[pdftex]{hyperref}
+\usepackage[algoruled, vlined,linesnumbered]{algorithm2e}
 
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmiccomment}[1]{\ \ // #1} % C-like // Comments
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows}
+\tikzstyle{block} = [draw, fill=blue!20, rectangle, 
+    minimum height=3em, minimum width=6em]
+\tikzstyle{sum} = [draw, fill=blue!20, circle, node distance=1cm]
+\tikzstyle{input} = [coordinate]
+\tikzstyle{output} = [coordinate]
+\tikzstyle{pinstyle} = [pin edge={to-,thin,black}]
+% \renewcommand{\algorithmicrequire}{\textbf{Input:}}
+% \renewcommand{\algorithmiccomment}[1]{\ \ // #1} % C-like // Comments
 
@@
 
+\newpage
 \section{Average Curve Computation}\label{chapter:Tracing}
+
+\begin{figure*}
+ \centering
+% The block diagram code is probably more verbose than necessary
+\begin{tikzpicture}[auto, node distance=2cm,>=latex']
+    \node [block, align=center] (traceCurve) {traceCurve\\\scriptsize Algorithm 2};
+    \node [block, right of=traceCurve, align=center, node distance=3cm] (snap) {snap\\\scriptsize Algorithm 3};
+    \draw [draw,->] (traceCurve) -- (snap);
+    \node [block, right of=snap, align=center, node distance=3cm] (findClosest) {findClosest};
+    \draw [->] (snap) -- (findClosest);
+    \node [block, right of=findClosest, align=center, node distance=3cm] (closestOnCubicHermite) {cubicHermite\\ \scriptsize Algorithm 1};
+    \draw [->] (findClosest) -- (closestOnCubicHermite);
+    \node [block, right of=closestOnCubicHermite, node distance=3cm] (solveQuintic) {solveQuintic};
+    \draw [->] (closestOnCubicHermite) -- (solveQuintic);
+    \node [block, right of=solveQuintic, node distance=3cm] (Laguerre) {Laguerre};
+    \draw [->] (solveQuintic) -- (Laguerre);
+    \node [block, below of=closestOnCubicHermite, align=center] (quadraticHermite) {quadraticHermite};
+    \draw [->] (findClosest) |- (quadraticHermite);
+    \node [block, right of=quadraticHermite, node distance=3cm] (solveCubic) {solveCubic};
+    \draw [->] (quadraticHermite) -- (solveCubic);
+\end{tikzpicture}
+\normalsize 
+\caption{\normalsize Block diagram for the major functions involved in tracing the average curve}
+\end{figure*}
 
@@
 TODO
+\blindtext
+
+\subsubsection{Newton-Raphson Method}
+TODO
+\blindtext
+
+\subsubsection{Laguerre's Method}
+TODO
+\blindtext
+\SetArgSty{textup}
+\begin{algorithm}[ht!]
+  \SetCommentSty{emph}
+  \KwIn{$\{P_0, T_0, P_1, T_1\}$: control points, $Q$: input point}
+  \KwResult{$t$: time in $[0,1]$ for the closest point}
+  $[a,b,c,d,e,f] \leftarrow $generateDistanceCoeffs$(P_0,T_0,P_1,T_1,Q)$; \\
+  $res \leftarrow $solveQuintic($a,b,c,d,e,f$); \\
+  $\{minDist, t^*\} \leftarrow \{\infty,-1\}$;\\
+  \ForEach{$r \in res$} {
+    \lIf{$|r.im| > 0$} \textbf{continue}; \\
+    \lIf{$r.re < 0\;$ \textbf{or} $\;r.re > 1$} \textbf{continue}; \\
+    $P_t \leftarrow \tau(P_0,T_0,P_1,T_1,r.re)$; \\
+    \If{$|P_t - Q| < minDist$} {
+      $\{minDist, t^*\} \leftarrow \{|P_t - Q|, r.re\}$;\\
+    }
+  }
+  \Return{$t^*$};
+  \caption{closestPointOnCubicHermite()}
+\end{algorithm}
 
@@
 TODO
+\blindtext
+
+\begin{algorithm}[ht!]
+  \SetCommentSty{emph}
+  \KwIn{$\{\mathbf{A},\mathbf{B}\}$: curves}
+  \KwResult{$\mathbf{C}$: average curve}
+  $\mathbf{C}.$add($\mathbf{A}[0], 0$); \\
+  $\{current, t_a, t_b, t_a^\prime, t_b^\prime\} \leftarrow \{\mathbf{C}[0],0,0,0,0\}$; \\
+  \While{true} {
+  $\vec{T}_A \leftarrow \vec{n}$($\mathbf{A}.\tau$($min(1,t_a+\epsilon)$) - $\mathbf{A}.\tau$($min(1,t_a)$)); \\
+  $\vec{T}_B \leftarrow \vec{n}$($\mathbf{B}.\tau$($min(1,t_b+\epsilon)$) - $\mathbf{B}.\tau$($min(1,t_b)$)); \\
+  $current \stackrel{+}\leftarrow \text{STEP\_SIZE} \;\vec{n}(\vec{T}_A + \vec{T}_B)$; \\
+  $\{t_a^\prime, t_b^\prime, r, P_c\} \leftarrow \text{snap}(\mathbf{A},\mathbf{B},current)$; \\
+  \lIf{$max(t_a,t_b) < 1$ \textbf{or} $t_a^\prime < t_a$ \textbf{or} $t_b^\prime < t_b$} \textbf{break}; \\
+  $\mathbf{C}.$add($P_c$); \\
+  }
+  \Return{$\mathbf{C}$};
+  \caption{traceCurve()}
+\end{algorithm}
+
+TODO
+\blindtext
+
+\begin{algorithm}[ht!]
+  \SetCommentSty{emph}
+  \KwIn{$\{\mathbf{A},\mathbf{B}\}$: curves, $Q:$ input point}
+  \KwResult{$\{t_a,t_b\}$: projections, $\{P_c, r\}$: sphere center/radius}
+  $P_c \leftarrow Q$; \\
+  \While{$i\texttt{++} < \;$MAX\_ITERS } {
+    $\{P_A, \vec{T}_A, r_A, t_A\} \leftarrow $closestPoint$(\mathbf{A}, P_c)$;  \\
+    $\{P_B, \vec{T}_B, r_B, t_B\} \leftarrow $closestPoint$(\mathbf{B}, P_c)$;  \\
+    $error \leftarrow |P_A - P_c| - |P_B - P_c|;$\\ 
+    \lIf{$error < \epsilon$} \textbf{break;} \\
+    $P_c \leftarrow P_c - error \;\vec{n}(\vec{T}_A - \vec{T}_B);$ \\
+    $\{d_A, \vec{V}_A\} \leftarrow $planeDist($P_c, P_A, P_B, \mathbf{A}.\tau(t_A + \epsilon))$;\\
+    $\{d_B, \vec{V}_B\} \leftarrow $planeDist($P_c, P_B, P_A, \mathbf{B}.\tau(t_B + \epsilon))$;\\
+    $P_c \leftarrow P_c - \text{PLANE\_PROJ\_COEFF} \;(d_A \vec{V}_A + d_B \vec{V}_B); $\\
+  }
+  \Return{$\{c_A.t, c_B.t, P_c, c_A.r\}$};
+  \caption{snap()}
+\end{algorithm}
 
@@
 
+\newpage
 \section{Visualization and Editing}\label{chapter:Visual}

### 2015-11-02 paper/paper.tex
Done with the text. Next: Feedback

@@
 
-\newpage
 \section{Average Curve Computation}\label{chapter:Tracing}
@@
 \subsection{Distance functions}
-TODO
-\blindtext
 
-\subsubsection{Newton-Raphson Method}
-TODO
-\blindtext
+To satisfy the distance constraints (1) and (2) in the objective function, we need to first 
+define the distance functions for a point and a curve. Note that since our curve representation
+is composed of multiple splines, we are actually interested in distance between a point
+and a hermite (quadratic or cubic). If $P(t)$ represents a point on the spline, parameterized
+with time $t$, we are interested in finding $t$ such that the tangent at that point $P^\prime(t)$ and
+the vector from that point to the input point $Q$ are perpendicular:
 
-\subsubsection{Laguerre's Method}
-TODO
-\blindtext
+\begin{equation}
+ P^\prime(t) \cdot (Q - P(t)) = 0
+\end{equation}
+
+If $P(t)$ is a quadratic polynomial in $t$, then the perpendicularity constraint is a cubic in $t$ and
+the points where the constraint are satisfied can be solved in closed form. Otherwise, if $P(t)$ is a 
+cubic polynomial in $t$, then the perpendicularity constraint is a quintic in $t$ and its roots
+need to determined with a local optimizer.
+
+Once all the roots of the perpendicularity constraint is found, the ones with imaginary coefficients
+and the ones with real values that are out of the [0,1] bound need to be eliminated. Moreover, to
+choose the closest perpendicular point, an iterative comparison has to be executed as shown in 
+Algorithm 1 for a quintic function. Note that the coefficients for the polynomial are obtained
+through Matlab's symbolic toolbox.
+
 \SetArgSty{textup}
@@
 
+\subsection{Optimization methods}
+
+Here, we would like to make a brief mention of the Java Commons Math library which provides a 
+plethora of optimizer and share our experience for future users.
+
+\subsubsection{Newton-Raphson Method}
+The Newton-Raphson method is one of the few methods that does not need an initial bracketing (limiting
+the solution space) and is also conceptually simple: just follow the derivative at the initial sample
+until it hits a zero-crossing and repeat until convergence. However, unfortunately, the library
+implementation does not always return the closest solution to the sample and thus, hard to use to 
+determine all the roots of the quintic function.
+
+\subsubsection{Laguerre's Method}
+The Laguerre's method is dependent on two initial bounds which have different signs when the 
+quintic is evaluated. The positive aspect of the implementation of the method is that it is guaranteed to find a solution
+if one exists within the bounds and if not, throw an exception. Thus, we discretize the
+time domain and search for the roots systematically.
+
 \subsection{Curve tracing}
-TODO
-\blindtext
+
+The block diagram that outlines the main functions in our curve tracing algorithm is presented in Figure 1.
+The main idea is to start with the intersection of the two curves as the first sample and continuously project
+out in the ``medial-line'' of the projected points' tangents iteratively. The traceCurve function in Algorithm 2
+demonstrates the steps involved.
 
@@
   \lIf{$max(t_a,t_b) < 1$ \textbf{or} $t_a^\prime < t_a$ \textbf{or} $t_b^\prime < t_b$} \textbf{break}; \\
+  $\{t_a, t_b\} \leftarrow \{t_a^\prime, t_b^\prime\};$ \\
   $\mathbf{C}.$add($P_c$); \\
@@
 
-TODO
-\blindtext
+For a sample \textit{current}, the algorithm first finds the tangents
+at the projected point $\vec{T}_A$ and $\vec{T}_B$, and \textit{projects out} the current sample by their median.
+However, since this point is not guaranteed to represent the distance constraints (1) and (2), nor since that
+sphere would be minimal, we need to incrementally move it with the \textit{snap} function.
+
+The pseudo-code for the \textit{snap} function is presented in Algorithm 3. At each iteration,
+two main changes are made on the initial sample. First, the point is moved towards its closest projections
+as shown in the lines 3-7. Then, the point is moved towards the planes defined by the projections 
+and the sphere center to minimize the radius of the sphere (in lines 8-10).
 
@@
 \subsection{Average Distance (geodesic) Sampling}
-TODO
 
-\subsection{Detect non-compatible control curves}
-TODO
+One challenge with the output of the naive tracing algorithm is that the samples on the curve $C$
+can have arbitrary relative distances. This can be an issue when additional visualization tools
+are built upon this curve such as tubes and boardwalk patterns. One approach is to predefine
+the number of points desired and sample the curve $C$ based on the total arc length and the 
+chosen resolution.
+
+Algorithm 4 outlines this approach where the relevant $C$ samples for the new sample are
+determined with a preliminary search (lines 6-12) and then, an intermediary point is computed
+with linear interpolation. Next, the interpolated point is pushed back on to the curve $C$
+to satisfy the object constraints 1-3 using the snap function in lines 14-15.
+
+The effect of the average distance sampling in comparison to the original average curve
+$C$ can be observed in Figure 3 between items (a) and (b). As expected, the samples
+along the geodesic curve is much more refined and can sample more thoroughly the ``incompatible''
+regions.
+
+\begin{algorithm}[ht!]
+  \SetCommentSty{emph}
+  \KwIn{$\{\mathbf{A},\mathbf{B}\}$: curves, $N:$ number of points}
+  \KwResult{$\mathbf{G}$: geodesic average curve}
+  $\mathbf{C} \leftarrow$ trace($\mathbf{A}, \mathbf{B}$); \\
+  $length \leftarrow \mathbf{A}.$arcLength()$ + \mathbf{B}.$arcLength(); \\
+  $\{\delta l, i, \mathbf{G}\} \leftarrow \{length / N, 0, \{\mathbf{C}[0]\}\}$;\\
+  \While{true} {
+    $\{g, \underline{i}, \underline{d}, \overline{i}, \overline{d}\} \leftarrow \{\mathbf{G}$.last, $i,0,-1,0\};$ \\
+    \While{$i < \mathbf{C}.$length} {
+      $d_A \leftarrow \mathbf{A}.$arcLength($proj(g,\mathbf{A}$), $proj(C[i], \mathbf{A}$)); \\
+      $d_B \leftarrow \mathbf{B}.$arcLength($proj(g,\mathbf{B}$), $proj(C[i], \mathbf{B}$)); \\
+      \lIf{$d_A + d_B < \delta l$} $\{\underline{i}, \underline{d}\} \leftarrow \{i, d_A + d_B\}$; \\
+      \Else {
+	$\{\underline{i}, \underline{d}\} \leftarrow \{i, d_A + d_B\};$ \\
+	\textbf{break};
+      }
+    }
+    \lIf{$i \geq \mathbf{C}.$length} break; \\
+    $P_c \leftarrow $LERP$(\mathbf{C}[\underline{i}], \mathbf{C}[\overline{i}], (\delta l - \underline{d}) / (\overline{d} - \underline{d}))$;\\
+    $\mathbf{G}$.add(snap($\mathbf{A}, \mathbf{B}, P_c$)); \\
+  }
+  \Return{$\mathbf{G}$};
+  \caption{geodesicTrace()}
+\end{algorithm}
+
+\newpage
+\subsection{Detect incompatible control curves}
+
+An important aspect of the average curve formulations provided in the first section is that the 
+there may not always be a solution to the three fundamental constraints of the average curve.
+In such cases, the average curve may not (1) exist, (2) be smooth, or (3) be close to the medial axis,
+and thus, it is important to inform the user when such cases may arrive as they manipulate the control points.
+
+One method of doing so is to monitor the arclength between the consecutive points in the average curve (geodesic)
+trace. If the arclength is significantly larger than the expected step size (1.5 times more in our implementation), 
+then a warning can be given by tracing the corresponding curve sections in red. Figure 2 below demonstrates
+examples of such extreme cases and the incompatible sections of the input control curves that would lead to a branching
+in the medial axis.
+
+\begin{figure}[ht!]
+  \includegraphics[width=1.0\linewidth]{images/Result45.png}
+	\caption{Extreme cases and non-compatibility detection}
+	\label{fig:Result4}
+\end{figure}
 
@@
 \end{figure*}
-\begin{figure*}
-	\centering
-		\includegraphics[scale=0.5]{images/Result4.png}
-		\includegraphics[scale=0.5]{images/Result5.png}
-	\caption{Example 3: extreme cases and non-compatibility detection}
-	\label{fig:Result4}
-\end{figure*}
 

### 2015-11-02 paper/paper.tex
Minor changes and suggestions

@@
 % \renewcommand{\algorithmiccomment}[1]{\ \ // #1} % C-like // Comments
+\usepackage{color}
+%\newcommand{}{\color{blue}}
+\newcommand{\cre}{\color{red}}
+\newcommand{\cbl}{\color{black}}
 
@@
 \title{Curve Average \\
-	{\large Project 3 for class CS6491 Computer Graphics}}
+	{\large Project 3 for CS 6491: Computer Graphics}}
 %\author{Sebastian Weiss}
@@
 \end{itemize}
-We now describe the single features in more detail
+We now describe the \cre single \cbl features in more detail.
 
@@
 
-As you can see, without geodesic sampling, the distances between the samples of the curve average are almost equispaced, but the closest projections on the control curve, indicated by the circular arcs, vary greatly in the distances.
+\cre As you can see, \cbl without geodesic sampling, the distances between the samples of the curve average are almost equispaced, but the closest projections on the control curve, indicated by the circular arcs, vary greatly in the distances.
 With geodesic sampling turned on, the curve average is sampled in that way that the sum of the distances between two consecutive closest projects on the control curves is constant.
@@
 \\
-We can now construct a unique circle in 3d with center $C$ and radius $r$ that goes through $A$ and $B$, the tangents at these points are $AP$ and $BP$ and lies in the convex hull of $A$,$B$ and $P$.
+We can now construct a unique \cre circle \cbl in $\mathbf{R}^3$ with center $C$ and radius $r$ that goes through $A$ and $B$, the tangents at these points are $AP$ and $BP$ and lies in the convex hull of $A$,$B$ and $P$.
 
@@
 
-We present a different approach now.
+\cre Create a new section?
+\cbl We present a different approach now.
 Instead of drawing circles orthogonal to the tangents, we directly take the vectors $P_i A_i$ and rotate them around $T_i$.

### 2015-11-02 paper/paper.tex
fixed some typos and doing some layout improvements

--- 2015-11-02 paper/paper.tex
@@
 \section{Objective}
-Formally, for two curves $A$ and $B$ defined by the control points $\{A_0 ... A_{n-1}\}$ and $\{B_0 ... B_{n-1}\}$, all points are in $\mathbb{R}^3$, and an interpolation scheme,
+Formally, for two curves $A$ and $B$ $\subset\mathbb{R}^3$ 
 the goal is to find a curve $C$ such that for each point $c \in C$, and its closest projections
@@
 For simplicity, we only focus on a special case that each control curve is defined by six control points $A_0$ to $A_5$ and $B_0$ to $B_5$, such that the first and the last control points are the same.
+We then interpolate between these control points to obtain the continuous curves $A$ and $B$.
 
@@
 
-Once all the roots of the perpendicularity constraint is found, the ones with imaginary coefficients
+Once all the roots of the perpendicularity constraint are found, the ones with imaginary coefficients
 and the ones with real values that are out of the [0,1] bound need to be eliminated. Moreover, to
@@
   \ForEach{$r \in res$} {
-    \lIf{$|r.im| > 0$} \textbf{continue}; \\
-    \lIf{$r.re < 0\;$ \textbf{or} $\;r.re > 1$} \textbf{continue}; \\
+    \lIf{$|r.im| > 0$} {
+			\textbf{continue}
+		}
+    \lIf{$r.re < 0\;$ \textbf{or} $\;r.re > 1$} {
+			\textbf{continue}
+		}
     $P_t \leftarrow \tau(P_0,T_0,P_1,T_1,r.re)$; \\
@@
   $\{t_a^\prime, t_b^\prime, r, P_c\} \leftarrow \text{snap}(\mathbf{A},\mathbf{B},current)$; \\
-  \lIf{$max(t_a,t_b) < 1$ \textbf{or} $t_a^\prime < t_a$ \textbf{or} $t_b^\prime < t_b$} \textbf{break}; \\
+  \lIf{$max(t_a,t_b) < 1$ \textbf{or} $t_a^\prime < t_a$ \textbf{or} $t_b^\prime < t_b$} {
+		\textbf{  break}
+	}
   $\{t_a, t_b\} \leftarrow \{t_a^\prime, t_b^\prime\};$ \\
@@
     $error \leftarrow |P_A - P_c| - |P_B - P_c|;$\\ 
-    \lIf{$error < \epsilon$} \textbf{break;} \\
+    \lIf{$error < \epsilon$} {\textbf{break}}
     $P_c \leftarrow P_c - error \;\vec{n}(\vec{T}_A - \vec{T}_B);$ \\
@@
 The effect of the average distance sampling in comparison to the original average curve
-$C$ can be observed in Figure 3 between items (a) and (b). As expected, the samples
+$C$ can be observed in Fig.\ref{fig:GeodesicSampling} between items (a) and (b). As expected, the samples
 along the geodesic curve is much more refined and can sample more thoroughly the ``incompatible''
@@
     \While{$i < \mathbf{C}.$length} {
-      $d_A \leftarrow \mathbf{A}.$arcLength($proj(g,\mathbf{A}$), $proj(C[i], \mathbf{A}$)); \\
-      $d_B \leftarrow \mathbf{B}.$arcLength($proj(g,\mathbf{B}$), $proj(C[i], \mathbf{B}$)); \\
+      $d_A \leftarrow \mathbf{A}.$arcLength($proj(g,\mathbf{A}), proj(C[i], \mathbf{A})$); \\
+      $d_B \leftarrow \mathbf{B}.$arcLength($proj(g,\mathbf{B}), proj(C[i], \mathbf{B})$); \\
       \lIf{$d_A + d_B < \delta l$} $\{\underline{i}, \underline{d}\} \leftarrow \{i, d_A + d_B\}$; \\
@@
     }
-    \lIf{$i \geq \mathbf{C}.$length} break; \\
+    \lIf{$i \geq \mathbf{C}.$length} {
+			\textbf{break}
+		}
     $P_c \leftarrow $LERP$(\mathbf{C}[\underline{i}], \mathbf{C}[\overline{i}], (\delta l - \underline{d}) / (\overline{d} - \underline{d}))$;\\
@@
 
-\newpage
-\section{Visualization and Editing}\label{chapter:Visual}
-
 \begin{figure*} %This just must appear at the page before the chapter "Displaying the average curve"
@@
 				& \includegraphics[scale=0.4]{images/ClosestProjections.png} & \includegraphics[scale=0.4]{images/CircularArcNet.png} \\
-			\small
-			a) no geodesic sampling & b) with geodesic sampling & c) the closest projections & d) net of circular arcs
+			\small a) no geodesic sampling & \small b) with geodesic sampling 
+			  & \small c) the closest projections & \small d) net of circular arcs
 		\end{tabular}
@@
 \end{figure*}
+
+\section{Visualization and Editing}\label{chapter:Visual}
 
@@
 \end{itemize}
-We now describe the single features in more detail
+We now describe the single features in more detail.
 
@@
 \\
-We can now construct a unique circle in 3d with center $C$ and radius $r$ that goes through $A$ and $B$, the tangents at these points are $AP$ and $BP$ and lies in the convex hull of $A$,$B$ and $P$.
+We can now construct a unique circle in $\mathbb{R}^3$ with center $C$ and radius $r$ that goes through $A$ and $B$, the tangents at these points are $AP$ and $BP$ and lies in the convex hull of $A$,$B$ and $P$.
 
@@
 However, this approach will lead to wrong results since the closest projections to the control curves are not orthogonal to the average curve tangent.
-Therefore, the distance from the closest projections to the average curve is not equal to the radius of the circle when it should touch the control curves.\\
+Therefore, the distance from the closest projections to the average curve is not equal to the radius of the circle when it should touch the control curves.
 
@@
   \centering
-	\begin{tikzpicture}[scale=0.5]
+	\begin{tikzpicture}[scale=0.6]
 		\draw[black,thick] (0,0) coordinate -- (0,10) coordinate; %curves
@@
 	\centering
-		\includegraphics[scale=0.4]{images/Result1.png}
-		\includegraphics[scale=0.4]{images/Result2.png}
+		\includegraphics[scale=0.5]{images/Result1.png}
+		\includegraphics[scale=0.5]{images/Result2.png}
 	\caption{Example 1: a bow}
@@
 	\centering
-		\includegraphics[scale=0.4]{images/Result3.png}
+		\includegraphics[scale=0.6]{images/Result3.png}
 	\caption{Example 2: a waterslide}
